---
title: "dplyr and pipes"
output:
  pdf_document: default
  html_notebook: default
---

# Install the tidyverse package set
Most usefull packages for modern R code : https://www.tidyverse.org/packages/ 
For those who learn to code R before 2015: http://kbroman.org/hipsteR/ 
Installs ggplot2, dplyr, tidyr, readr, purr, tibble and much more:
‘colorspace’, ‘mnormt’, ‘bindr’, ‘RColorBrewer’, ‘dichromat’, ‘munsell’, ‘labeling’, ‘viridisLite’, ‘rematch’, ‘plyr’, ‘psych’, ‘reshape2’, ‘assertthat’, ‘bindrcpp’, ‘glue’, ‘pkgconfig’, ‘rlang’, ‘R6’, ‘Rcpp’, ‘BH’, ‘plogr’, ‘digest’, ‘gtable’, ‘scales’, ‘lazyeval’, ‘mime’, ‘curl’, ‘openssl’, ‘cellranger’, ‘stringi’, ‘selectr’, ‘tidyselect’, ‘broom’, ‘dplyr’, ‘forcats’, ‘ggplot2’, ‘haven’, ‘httr’, ‘hms’, ‘jsonlite’, ‘lubridate’, ‘magrittr’, ‘modelr’, ‘purrr’, ‘readr’, ‘readxl’, ‘stringr’, ‘tibble’, ‘rvest’, ‘tidyr’, ‘xml2’

```{r, warning=F, message=F}
# install.packages("tidyverse")
library(tidyverse)
library(magrittr)


# other useful depedencies
library(tableone)
library(psych)
library(fBasics)


# Sometimes function mask each other !
# tib %>% filter(Sex == "Male", Eye=="Brown") # produce error
filter = dplyr::filter 
# tib %>% filter(Sex == "Male", Eye=="Brown") # now its ok


# vignette("dplyr")
```

# Load some datesets
```{r}
#install.packages("datasets")
# install.packages("mlbench") 

# library(help= "mlbench") (Glass, BreastCancer, BostonHousin, Ionosphere, PimaIndiansDiabetes, Sonar)
library(datasets)
library(mlbench)
data("BostonHousing")
head(BostonHousing)

data("HairEyeColor")
HairEyeColor

data("mtcars")
head(mtcars)

data("swiss")
head(swiss)

data("BreastCancer")
head(BreastCancer)


```


# Manipulate data with dplyr and the pipe operator
Add a pipe operator : CTRL + SHIFT + M

```{r} 
tib = as.tibble(HairEyeColor)

# Filtering and selecting

filter(tib, Sex== "Male")

tib %>% 
  filter(Sex == "Male", Eye=="Brown") %>% 
  select(Eye)

# library(data.table)
# dt = data.table(tib)
# dt[ Sex == "Male" & Eye=="Brown", Eye, with=F]

# LHS %>% RHS == RHS(LHS)

select(filter(tib, Hair == "Red"), Eye) 

```


```{r}
tib %>% select(contains('S'))
tib %>% select(Eye:n)

tib %>% select(-Eye)



mtcars %>%  
  select(mpg, cyl, wt, am) %>% 
  group_by(cyl) %>%
  summarise(avgmpg = mean(mpg), avgwt = mean(wt)) %>%
  filter(avgmpg > 20)

# cw %>% group_by(sid) %>% summarise(accuracy = mean(correct)) %>% arrange(desc(accuracy))

# select(iris, contains("."))
# Select columns whose name contains a character string.
# select(iris, ends_with("Length"))
# Select columns whose name ends with a character string.
# select(iris, everything()) Select every column.
# select(iris, matches(".t."))
# Select columns whose name matches a regular expression.
# select(iris, num_range("x", 1:5)) Select columns named x1, x2, x3, x4, x5.
# select(iris, one_of(c("Species", "Genus")))
# Select columns whose names are in a group of names.
# select(iris, starts_with("Sepal"))
# Select columns whose name starts with a character string.
# select(iris, Sepal.Length:Petal.Width)
# Select all columns between Sepal.Length and Petal.Width (inclusive).
# select(iris, -Species)
# Select all columns except Species.

```

## Assignment

```{r}

df = mpg %>% filter(cty > 20, cyl == 4) %>% select(1:2)

mpg %>% filter(cty > 20, cyl == 4) %>% select(1:2) -> df

df %<>% filter(manufacturer != "honda")   

df %>% table %>% as.tibble %>% filter(n!=0)


all_letters <- c(letters, LETTERS) %>%
    sort %>% 
    write.csv(file = "all_letters.csv")


```

## Using functional sequences

```{r}
library(magrittr) # needed to include the pipe operators
library(lubridate) # in tidyverse 
read_year <- . %>% as.character %>% as.Date %>% year

# Creating a dataset
df <- data.frame(now = "2015-11-11", before = "2012-01-01")
#          now     before
# 1 2015-11-11 2012-01-01

# Example 1: applying `read_year` to a single character-vector
df$now %>% read_year
df$now %>% as.character %>% as.Date %>% year
# [1] 2015

# Example 2: applying `read_year` to all columns of `df`
df %>% lapply(read_year) %>% as.data.frame  # implicit `lapply(df, read_year)
#    now before
# 1 2015   2012

# Example 3: same as above using `mutate_all`
library(dplyr)

a = c("albert happy", "albert sad", "albert writing")
replace_me = function (data) { gsub(pattern="albert ", replacement = "Kengo ", x = data) }

replace_me(a)

data.frame(name_emotion = a) %>% mutate_all(funs(replace_me))

df %>% mutate_all(funs(read_year))
# if an older version of dplyr use `mutate_each`
#    now before
# 1 2015   2012
```


# Plots
```{r}
library(ggplot2)


data("mtcars")
mtcars$name = rownames(mtcars)
mtcars$name = NULL

mcars = mtcars %>% mutate(name = rownames(.)) %>% select(name, everything())

#select(mcars, name, everything())

ggplot(data= mcars, aes(x=cyl, y=mpg, col=name)) + geom_point()

diamonds %>% 
    filter(depth > 60) %>% 
    group_by(cut) %>% 
    summarize(mean_price = mean(price)) %>% 
    ggplot(., aes(x = cut, y = mean_price)) + 
        geom_bar(stat = "identity")
```

# Basic renaming
```{r, results='asis'}
require(pander)

# example of variable renaming
breastcancer = BreastCancer %>% rename_all(tolower) 
cat("toLower", pandoc.list(names(breastcancer)))

# you could also use base functions
breastcancer %<>% setNames(tolower(gsub("\\.", "_", names(.))))
cat("setNames(tolower(gsub(.)))", pandoc.list(names(breastcancer)))
```

# Reshaping 
Base R function for reshaping datasets is reshape
Some very usefull parameters
```{r}
# Base R function
class(Indometh)
wide <- reshape(Indometh, v.names = "conc", idvar = "Subject",
                timevar = "time", direction = "wide")

df <- data.frame(id = rep(1:4, rep(2,4)),
                 visit = rep(c("Before","After"), 4),
                 x = rnorm(4), y = runif(4))
reshape(df, timevar = "visit", idvar = "id", direction = "wide")
reshape(df, timevar = "visit", idvar = "id", direction = "wide", v.names = "x")
 
df3 <- data.frame(id = 1:4, 
                  age = c(40,50,60,50), 
                  dose1 = c(1,2,1,2),
                  dose2 = c(2,1,2,1), 
                  dose4 = c(3,3,3,3))

reshape(df3, direction = "long", varying = 3:5, sep = "")

```

# Use TidyR / Reshape2
Best organization for datasets

Rules:

  - Each variable has its own column
  - Each observation in its own row
  - Each value is placed in its own cell 

Functions gather, spread, separate, and unite are the building block of the tidyR library
Here parameters are very simple so you have to have a good understanding of how you want to reshape your data. 

Sometimes what you really want to do is create new variables: for that you need all the smaller functions to make it longer first, unite variables, 
then spread it out so that each variable has its own column.
```{r}
stocks = data.frame(
  time = as.Date('2009-01-01') + 0:9,
  X = rnorm(10, 0, 1),
  Y = rnorm(10, 0, 2),
  Z = rnorm(10, 0, 4)
)
head(stocks)

# Gather all the selected variables in two columns stock and price
# you can use the minus sign to exclude columns from the selection
stocksm = stocks %>% gather(stock, price, -time)
stocksm %>% head

# You can invert what you just did using spread
stocksm %>% spread(stock, price) %>% head

# But you can also spread based on another variable
stocksm %>% spread(time, price) %>% head


# Use 'convert = TRUE' to produce variables of mixed type
df <- data.frame(id = rep(c(1, 2), each = 3),
                 var = c("Sepal.Length", "Species", "Species_num"),
                 value = c(5.1, "setosa", 1, 7.0, "versicolor", 2))
df
df %>% spread(var, value) 
df %>% spread(var, value, convert = TRUE) 

  
# Unite several columns into one 
storms %>% unite(date, day:year, sep = "-")


# Separate values into new column based on pattern
example <- tibble(`N [ears]` = c("173", "60", "54 [96]", "168 [328]", "906 [1685]"), 
                  `% Otorrhea` = c("58.61%", "13.30%", "11.11%", "52.38%", "14.79% [10.45%]"))

example %>% 
    separate(`N [ears]`, into = c("N_patients", "N_ears"), sep = " \\[", fill = "right") %>%
    separate(`% Otorrhea`, into = c("pct_patients", "pct_ears"), sep = "\\s\\[", fill = "right") %>%
    mutate_each(funs(parse_number))

```

```{r}
# Exercice 
# We will try to do what the reshape function did for us on a 
df <- data.frame(id = rep(1:4, rep(2,4)),
                 visit = rep(c("Before","After"), 4),
                 x = rnorm(4), y = runif(4))
reshape(df, timevar = "visit", idvar = "id", direction = "wide")

```


```{r}






# SOLUTION

 # first gather variales into two columns called variable and value 
df %<>% gather(variable, value, -(c("id", "visit"))) %>% print

# Then create the new variables X before X after and Y before Y after by uniting the two columns
df %<>% unite(variable, variable, visit, sep = "_") %>% print

# Finally spread into one column per variable
df %<>% spread(variable, value) %>% print
  
```


melt and cast are the two main functions of reshape2
```{r}
# convert matrix to data.frame
HairEyeColor %>% reshape2::melt() -> tib
tib = as.tibble(HairEyeColor) # !! Does not transforms character columns as factor

```


# Basic stats 
```{r}
# Generate hypothesis 
# Generate data using sampling
# Y dependent
# A B C D E F G independents continuous / discrete
# linear model 
Y = A*b1 + intercept + e

rmultinom(10, size = 12, prob = c(0.1,0.2,0.8))
A = 1:10
Y = 2*A + e

plot(Y,A)

# heights depends on grades or is predicted by grades 
heights = rnorm(n = 40, mean = 175, sd = 15)
grades =  rnorm(n = 40, mean= 3, sd = 2)

happyness = (1/4)*grades + 1*heights + rnorm(40, 0, 400)

heights.n = scale(heights)
grades.n  = scale(grades)
happyness.n = scale(happyness)
summary(lm(happyness ~ grades + heights)) 
mfull = lm(happyness ~ grades + heights)
summary(mfull)
mrestricted = lm(happyness ~ 1)
summary(mrestricted)
anova(mfull, mrestricted)
summary(lm(happyness.n ~ heights.n )) 

newSubject = data.frame(grades = rnorm(n = 40, mean= 3, sd = 2), heights = rnorm(n = 40, mean= 160, sd = 12)) 
predict(m, newdata = newSubject)

```
```{r}
?distribution

leftToLive = intercept + b1*G(0 young 1 old) + b2 *G*cigarets 

# regression
lm(formula = happyness ~ grades + heights + grades:heights)
lm(formula = happyness ~ grades * heights )
lm(formula = happyness ~ grades/heights)
lm(formula = happyness ~ grades + grades:heights)
lm(formula = happyness ~ grades * heights * var4 * var5 - var4:var5)
lm(formula = happyness ~ grades * heights * var4 * var5 - var4:var5)
lm(formula = happyness ~ 0 + grades^2 +grades + I(heights * grades))
aov(formula = ,data = )
grades^2 + 2*grades*height + heights^2

happyness = intercept + b1*grades + b2*heights + e

b1 = -30
# ?distribution
# Main vs Interaction vs Nested effect 
# Regression vs Anova
# Linear regression 
# Logistic regression 

```

# Pandoc 

```{r, results='asis'}
require(fBasics)
mpg %>% mutate_all(as.numeric) %>% basicStats %>% t %>% pandoc.table
```


# df2latex
A couple of xx2latex function allows to transform data into a latex format that 
can help produce clean graphics. To show the graphics or in general to use the output
of a code chunk as if you wrote the RMarkdown code directly, use the chunk option results="asis"
```{r, results="asis", warning=F}
require(fBasics)
require(psych)

# some utilities
to.latex.table = function (data, ...) {
    data %>% 
        df2latex(silent = T, file="tempcorrtest.txt", ...) %>% 
        gsub(pattern = "\\begin{tabular}", replacement = "\\resizebox{\\textwidth}{!}{\\begin{tabular}", ., fixed = T) %>% 
        gsub(pattern = "\\end{tabular}", replacement = "\\end{tabular}}", ., fixed = T) %>%  
        paste0(collapse = " ") 
}

corr.test.2latex = function (data, variables = NULL, method="spearman", ...) {
    if (is.null(variables)) { variables = names(data)}
    
    rs = corr.test(data %>% select(variables), method=method) #find the probabilities of the correlations
    cp =corr.p(r=rs$r, n = NROW(data))
    
    cpp = rs$r
    cpp[cp$p>0.05] = 3
    
    colnames(cpp) = paste0(1:NROW(cpp), ".") 
    rownames(cpp) = paste(colnames(cpp), gsub("_", " ", rownames(cpp), perl=T))
     
    cpp[upper.tri(cpp)] = NA
    cpp %>%
        df2latex(silent = T, file="tempcorrtest.txt", ...) %>%
        gsub(pattern = "3.00",replacement = ".", ., fixed = T) %>%
        gsub(pattern = "\\begin{tabular}", replacement = "\\resizebox{\\textwidth}{!}{\\begin{tabular}", ., fixed = T) %>% 
        gsub(pattern = "\\end{tabular}", replacement = "\\end{tabular}}", ., fixed = T) %>%  
        paste0(collapse = " ") 
    
}

shapiro.test.p = function (x) { shapiro.test(x)$p.value }
shapiro.test.star = function (x) {if (shapiro.test(x)$p.value<0.05) { stars.pval(shapiro.test(x)$p.value) } else { "." } }

cat(mpg %>% mutate_all(as.numeric) %>% basicStats %>% t %>% to.latex.table)


cat(breastcancer %>% select(-id) %>% mutate_all(as.numeric) %>% corr.test.2latex)

```

# Tables
A useful package to create tables is tableone.
install.packages("tableone")
```{r}

library(tableone)
generate_table = function(data, columns = NULL, strata_g = NULL) {
  if (is.null(columns)) { columns = names(data) }
  columns = unique(columns)
  catVar = NULL
  sapply(columns, function(name) {
    if (class(data[[name]])%in%"factor") {
      catVar <<- c(catVar, name)
    }
  })

  data_filtered = data %>% select(unique(c(columns,strata_g)))

  if (is.null(strata_g)) {
    table_g <- CreateTableOne(vars = columns, data = data_filtered, factorVars = catVar, includeNA = T)
  } else {
    table_g <- CreateTableOne(columns, data_filtered, catVar, strata = strata_g, includeNA = T)
  }


  mat_g <- print(table_g, exact = "stage", quote = FALSE, noSpaces = TRUE, printToggle = FALSE)

  return(mat_g)
}

breastcancer %<>% mutate_each(funs(as.numeric), -id, -class) 
breastcancer %>% select(-id) %>% generate_table(strata_g = "class")


```



## Recode
http://dplyr.tidyverse.org/reference/recode.html
```{r}
x <- sample(c("a", "b", "c"), 10, replace = TRUE)
recode(x, a = "Apple")
```


# TidyR
```{r}

 

# nested dataframes
# data %>% group_by(x) %>% nest()
```


# BROOM
http://varianceexplained.org/r/broom-intro/
```{r}
```


# Survival
https://github.com/pavopax/gists/blob/master/survival-in-R.md 

# Purrr

```{r}
 

```
