---
title: "introduction"
author: "Albert Buchard"
date: "12/20/2016"
output: pdf_document
--- 
# How to keep your code clean
## Coding convention
* Pick a naming convention and stick to it
  + camelCase = "this is a nice style"
  + snake_case = "this is ok too"
* Comment your code
* Look at the google style book to make sure your code is easilly readable by anyone
  + https://google.github.io/styleguide/Rguide.xml 
  + they advice to use only "<-" and not "=" but i personally think it is pointless
  
## Storage
Keep a README.md file at the root of your folder explaining where everything is, 
helping someone that knows nothing about your data to navigate your work.
Keeping your work in the cloud, through services like dropbox, icloud, or google drive. 
The best would be github but it is not easy in the begining.

### Folders
Keep your folder clean, with clear names in minuscules separated by "_" :

* data
    + raw
    + preprocessed
    + analysis
        + analysis_one ...
        
* scripts
    + preprocessing: scripts that transforms the raw data in processed data
    + analysis: scripts that use preprocessed data and performs analysis on it
    + markdown: your markdown files
    + r_files: other R files, like utility functions
    
* media: here should go any ressources, presentations, images you produced or needed etc...
    + presentations
    + graphics
    + text
    + notes
    
* backups: you might need a backup folder when in doubt
    + data
    + script
    + media
    
# Variables 

R infer on its own the type of the variable that you want to create based on the input you give.
All variables are at minimum a vector.

### Atomic vector data type
```{r atomic vectors}
# Character
a <- "This is a character vector"

# Numeric (integer)
a = 12
a <- 12

# Numeric (Float)
a <- 12.2

# Logical
a <- TRUE
a <- FALSE

print(a)
```


### Combine atomic elements
#### Vectors needs to be of same type
```{r combine vectors}
a <- c(1,2,3)
print(a)

a <- c(1,"a")
print(a)
```

#### List can mix types
```{r}
a <- list(1,2,3)
print(a)

a <- list(1,"a")
print(a)
```

### Matrix 2*2
Needs to be of same type
```{r}
a <- matrix( c('a','a','b','c','b',2), nrow = 2, ncol = 3, byrow = T)
print(a)

a <- matrix( c('a','a','b','c','b',2), nrow = 2, ncol = 3, byrow = F)
print(a)
```

### Array N*N
Needs to be of same type
```{r}
a <- array(c('green','yellow'),dim = c(3,3,2))
print(a)
```

### Factor
For categorical variables
```{r}
# Create a vector.
apple_colors <- c('green','green','yellow','red','red','red','green')

# Create a factor object.
factor_apple <- factor(apple_colors)

# Print the factor.
print(factor_apple)
print(nlevels(factor_apple))

# Change names of the factors
levels(factor_apple) <- c("Kindof Green", "Kindof Red", "Kindof Yellow")
print(factor_apple)
```

### Dataframe ++
```{r}
first_names = c("Melissa", "Sibylle", "Zoe", "Maria")
ages = c(23, 22, 24, 25)

df <- data.frame(first_name = first_names,
                 age = ages,
                 subject = as.character(c("Activity", "Motivation", "Fluid Intelligence", NA)))



print(df)
print(df$age)
```
### Missing values are "NOT ASSIGNED" or "NA"
```{r}
print(df$subject)
print(is.na(df$subject))
```

### I dont want a factor, I want characters !
Sometimes you have ot use function such as apply or sapply, that performs simple loops on your data.
```{r}
print(sapply(df[, 3], as.character))

df[, 3] <- sapply(df[, 3], as.character)

print(df)
```

# Operators
## Relational operators
```{r}
print(12>23)

a = 12
print(a == 12)

print(a != 32)

print(a >= 11)

a = 32
```

## Tests
```{r}
if (a == 432) {
  print("a est egale a 12 !")
} else {
  print("pas egale a 12")
}

```

## Logical operators
```{r}
print((12>23)&&(12<23))

a <- 12
print((a>20)||(a==12))

print(!(a == 32))

!is.na(a)

```

### Element wise logic
When a vector is tested against a vector of same length
```{r}
a <- c(F, T, T)
b <- c(T, F, T)
print(a&b)
print(a|b)

```

### Others
```{r}
a <- 1:8
print(a)

a <- rep("ce qui est repete", 4)
print(a)
```

# Flow control statements
Flow controls statements are all thestatement of a language that will redirect the flow of execution of a program. 

## Conditional control
Sometime you want to execute something only if a condition is true.
The most used is the "if/else if/else" statement. 
```{r}
a = c(F,F,F,T)

if (a[1]) {
  print("first")
} else if (a[2]) {
  print("second")
}  else if (a[3]) {
  print("third")
}  else if (a[4]) {
  print("quatrieme")
} else {
  print("invalid")
}

b <- c(F,F,F,T)

# TODO: write a statement that checks if b has any of its value equal to TRUE. 
# If it does return all the indices of the TRUE values
# If not, say that you did not find a True value in any of the %SIZE% elements of b
# (hint:: ?any and ?which)

```

When you only want to check the value of _ONE_ variable. Another way is to use the _switch_ statement. It test the value of a variable against several possibilities, like so:
```{r}
strangeName <- "Grabulas"
switch(strangeName,
       "BJ Gabbour" = {
         print("It was Gabbour all along !")
       },
       "Hortiche" = {
         print("She's just everywhere")
       },
       "Grabulas" = {
         print("Run you fools !")
       })
```


## Loops
You often need to repeat some statement. That's what _for_ and _while_ are here for !
```{r}

for (i in 1:NROW(df)) {
  person = df[i,]
  
  print(paste0(person$first_name, person$age))
}
```

```{r}
i = 0
while(i<10) {
  print(i)
  i = i + 1
}

```


## Functions
Functions are the central concept to programming. You can think of it as a box containing a series of instructions. Usually they take input, perform change on it, and returns a value. 
Note that they do not always take input or return a value and act only a series on instructions that do not takes input and do not return anything, but for example will read or write some information on the disk, or setting some parameter.

### A simple sum
```{r}
sommeFunction = function  (x, y, z) {
  # Do some action on the parameters
  sum = x+y+z
  
  # Return to sender the result of your computation
  return(sum)
}

somme = sommeFunction(1,2,3)
```

But of course R has a better function already built in !
```{r}

sum(c(1,2,3))

```

### A function that generates participant ids
What if you want to set some default parameters ?
Here is a more complex function that plays with strings to create random IDs for your subjects.
```{r}
getRandomId = function(numberOfIds = 1, lenght=12, allowedCharacters = c(0:9, letters, LETTERS))
{
    # initialize vector
    randomStrings = c(1:numberOfIds)
    
    # start the generation loop
    for (i in 1:numberOfIds)
    {
        randomStrings[i] <- paste(sample(allowedCharacters, lenght, replace=TRUE), 
                                  collapse="")
    }
    
    # return the strings
    return(randomStrings)
}

# TODO Now generate 650 ids !
# ids = ...

# Add the ids to the data
```

### Create your own function
Choose one between those three possible function, and create them:
* A function that returns the product of two numbers such that aTimesb = product(a,b)
* A function that adds a prefix to a string, such that prefixedString = prefix(prefixString, string)
* A function that takes out the mean of each column of a data frame, and divides by the standard deviation (process called normalization)

# Libraries
## Install libraries
```{r, eval=F}
install.packages("ggplot")
install.packages("psych")
```

## Load libraries
```{r}
library(ggplot2)
library(psych)

#help(package= "psych")
#vignette("ggplot2-specs", package = "ggplot2")
```

# Explore your data
## Load data
Usually you will load three type of data:
* Excel files: .xslx
* Comma separated values: .csv
* R data file: .RData

You load them differently
```{r}
# Load a RData
setwd("~/Google Drive/Master Students/courses/introduction_a_r")
load("data/raw/data.RData")
#or if you want to rename your data
renamedData = get(load("data/raw/data.RData"))
```

## When in doubt, google it !
```{r}
# But.. how to load EXCEL FILES ?
# TODO Check stackoverflow / Google and load excel and csv file
# "load xlsx file R"
# "load csv file in R"
# can you copy paste ?? read.clipboard 
```

## Explore your data
Looking at your data before starting asking question is important to detect errors you might have made, wrong IDs, numbers of NA, wacky values...
https://cran.r-project.org/web/packages/psych/vignettes/overview.pdf 
```{r, warning=F}
# View(data)

# Psych has a lot of tools for exploratory analysis
library(psych)
describe(data)
describeBy(data, group = "isOlder")
describeData(data)
head(data)

# Some quick plots
error.bars(data[, c("isOlder", "performance")])
boxplot(data[, c("performance")])
pairs.panels(data[-6])
corPlot(data.matrix(data[-6]))


# Look for outliers
# outlier(data[, c("performance", "IQ", "responseTime")],cex=.8)

# Filter your data
dataScrubbed = scrub(data,2:3,min=c(88,1300), max=c(115, 1600), newvalue=NA)
old = subset(data,data$isOlder==1)
scatter.smooth(old$age, old$performance)
```

### Correlations
```{r}
# What is a correlation ?
draw.cor(expand=20,cuts=c(0,0),r = 0.57)
```

Plot Performance as a function of IQ. Performance = f(IQ).
```{r}
ggplot(data, aes(x=IQ, y=performance)) + geom_point()
```

Now, what kind of relation do you think exist between the two? 
```{r}
ggplot(data, aes(x=IQ, y=performance)) +
    geom_point(shape=1) +    # Use hollow circles
    geom_smooth(method=lm,   # Add linear regression line
                se=FALSE)
```

Usually you want to know the interaction between multiple variables. For example the performance might be linked to the IQ, the Age, and/or the university !
```{r}
# regression 
regression = lm(formula = performance ~ IQ + age + university + IQ:age, data = data, na.action = na.omit)
summary(regression)
data$prediction = 3.497e-03 * data$IQ + 2.162e-01 * data$university 
ggplot(data, aes(x=IQ, y=performance)) +
    geom_point(shape=1) +    # Use hollow circles
    geom_smooth(method=lm,   # Add linear regression line
                 se=FALSE)+
    geom_point(aes(x=IQ, y=prediction), shape=17)+ 
    geom_line(aes(x=IQ, y=prediction), color="red") 

ggplot(data[1:20,], aes(x=IQ, y=performance, color=id)) +
    geom_point(shape=1)+    # Use hollow circles
    geom_smooth(method=lm,   # Add linear regression line
                 se=FALSE, color="blue")+
    geom_point(aes(x=IQ, y=prediction, color=id), shape=17)+ theme(legend.position="none")
  

# But is this model the good one ?
regression2 = lm(formula = performance ~ IQ * age , data = data, na.action = na.omit)
summary(regression2)

# Anova(model1,model2,test="Chisq") allows to compare models
anova(regression,regression2,test="Chisq")
```

```{r}
# Other way to analyses correlation (usually through covariance matrix)
# correlation and analysis
corr.test(data.matrix(data[-6]))
corPlot(data.matrix(data[-6]))

# Multivariate correlation predict y columns with x
setCor(y = 1:4,x=c(5,7),data=data)

```

# Mediation analysis
```{r}
mediate(y = 4, x = c(5,7), m = 2, data = data)
```

# Dimensionality reduction
```{r}
# How many components should you expect ?
dataWithoutId = data[-6]
fa.parallel(dataWithoutId)

# Data loads on variable 
principal(dataWithoutId, nfactors = 2)

# Latent variable loads on data (age, IQ, university)
factanal(dataWithoutId, factors = 3)
fa.diagram(fa(dataWithoutId, nfactors = 3))

# Hierarchical
#install.packages("GPArotation")
library(GPArotation)
omega(dataWithoutId)

#Clusters
iclust(dataWithoutId)
```

```{r}
# structural equatiom modelin
sem = esem(r = cor(dataWithoutId), varsX = c(5,6), varsY = 1:4, nfX = 2, nfY = 1,
n.obs = 650, plot = FALSE)
print(sem)
esem.diagram(sem)
```


# How to generate fake data
```{r}
# Number of subjects
N = 650

# Sample from a list
age = sample(18:65, N, replace = T)

hist(age)
```

```{r}
# Sample from gaussian distribution to generate IQ
baseIq = rnorm(N, 100, 15)
agedIq = baseIq * (log(18)/log(age/2))
data = data.frame(age = age, isOlder = age>60, IQ = agedIq) 

plot(age, (log(18)/log(age/2)))
plot(age,agedIq)


# A slightly more beatifull way of plotting 
# install.packages("ggplot")
library(ggplot2)

# Line plot + prediction error
ggplot(data, aes(x = age, y = IQ, color=age)) + geom_point() + geom_smooth()

```

```{r}
# Histograms
ggplot(data, aes(IQ, ..density.., fill = isOlder)) +
geom_density(alpha=0.2)  

```

```{r}
# Sample from a custom distribution
# For example the Exponential 
# P(x) = lambda * exp (-lambda*x)
# lambda = 1 / mean
# CDF = 1 - exp(-lambda*x) ==> x = - ln (1-CDF) / lambda
mean = 1500
lambda = 1 / mean
CDF = sample(0:100, N, replace = T) / 100
x = -log(1-CDF) / lambda
```

### Complex distribution
Sample from from complex distribution such as Diffusion Model for Response Time
Lets image you are slower with age !
```{r}
# Here bias depends on age
responseTime = rep(NA, N)
currentAccumulators = rep(0,N)
bias = sample(c(-1,1), N, T)+rnorm(N,0,0.1)
threshold = 1000
for(i in 1:5000) {
  currentAccumulators = currentAccumulators + bias * abs(rnorm(N, 0,1-(0.008*age)))
  indices = which(abs(currentAccumulators) > threshold)
  newIndices = indices[!indices%in%which(!is.na(responseTime))] 
  responseTime[newIndices] = i
}

hist(responseTime)
data$responseTime = responseTime

ggplot(data, aes(age, responseTime)) + geom_smooth(alpha=0.2) + geom_point()
ggplot(data, aes(responseTime, ..density.., fill=isOlder)) + geom_histogram(alpha=0.5)

```

### Create fake correlation in the data
For example imagine IQ is linked to performance and wether or not they went to university
```{r}
# University
university = rep(0,N)
university[which(rnorm(N)>-0.25)] = 1

# Generate fake performance linked to IQ
performance = rnorm(N,60,15)*agedIq/100 + 30*university

# Scale between 0 and 1
performance = performance - min(performance)
performance = performance / max(performance)

data$performance = performance
data$id = getRandomId(N)
data$age = age
data$university =university
```
# Save your data
```{r}
# Set your working directory - your reference for the file system
setwd("~/Google Drive/Master Students/courses/introduction_a_r")

# Save into your data/raw
save(data, file = "data/raw/data.RData")

# CSV format
write.csv(data, file = "data/raw/data.csv", row.names = FALSE)

# XLS - needs a library
library(WriteXLS)
WriteXLS(data, "data/raw/data.xls")
```

# Example: Scaling function
```{r}
scale_by <- function (data, by = "minmax") {
        copied_data = data.frame(data)
        columns = names(data)
        
        switch(by,
               meanvar = {
                       center <- mean
                       spread <- sd
               },
               medianvar = {
                       center <- median
                       spread <- sd
               },
               minmax = {
                       center <- min
                       spread <- max
               })
        
        for (column in columns) {
                center_by = center(copied_data[[column]], na.rm = T)
                reduced_by = spread(copied_data[[column]], na.rm = T)
                copied_data[[column]] = (copied_data[[column]] - center_by) / reduced_by
        }
        return(copied_data)
}
```

# Debugging
A good rule of thumb is that you will introduce a bug in your code every ten new lines.
Besides real design issue in your code, bugs are usually due to overlooking certain
extreme "use case" that you did not plan for, or because you did not pay attention
to the values of your variables and content of your data.

Quick fix bugs:
- Wrong variable name
- Forgot a comma or parenthesis

Harder bugs:
- Wrong format
  + Using a factor as a string or numeric
  + Some NA in the data
  + Some misformated strings or outlier values
- Wrong data size
- Silent bugs - return incorect values (+/- design issues)

Three steps to correct a bug:
- Reproduce
- Isolate : harder part
- Correct
  
```{r}

# Break on error
# browser()
# breakpoints
# print


```
