---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

# A bit of stats

Does eating a banana makes you happier ?
Lets assume happiness is distributed following a gaussian distribution.
You have a way of measuring it reliably.

- You are alone eating a banana: 
    + what is the magnitude ?
    + what is the significance ?
- You are doing it with 4 friends
    + what is the magnitude ?
    + what is the significance ?
- You are doing it with 40000 people 
    + Let's say you see a 3% increase in happiness
    + Would it be significant ?
- You are doing it with 40000 in china and 40000 in the US
    + Will the effect be significantly different ?
    + How will you know if that difference is really meaningful ?

```{r}
before_happiness = rnorm(40000,7,5)
after_happiness = before_happiness + rnorm(40000,0.5,1)
mean(after_happiness - before_happiness)

# What about 
after_happiness = before_happiness + rnorm(40000,0.5,35)
```

# Power Analysis

A concrete example, you want to know how many subject are necessary for your comparative experiment.
You have two groups:
  - Group 1 is doing A then B
  - Group 2 is doing B then A
  
Research questions: 
  - Do you need groups to test if task B is correlated to task A ?
  - You are wondering how many subject are necessary to show a difference in performance between the two groups on task A and B
  - Is this difference larger for task A or task B. 
  - How many subject would you need if you wanted to know the exact magnitude of the effect size with precision ?
```{r}
# Comonly used values in clinical trial design
qnorm(0.05)
pnorm(1.644854)
power = 0.9
false_positive_error_rate = 0.05
alpha = false_positive_error_rate

# Specific to the problem
effect_size = 0.1
standard_deviation = 0.22
variance = standard_deviation^2

mean = 0.64
estimated_n = 364
n = estimated_n

quantile = qnorm(0.05, mean, standard_deviation)-(effect_size/(standard_deviation/sqrt(n)))
estimated_power = 1 - pnorm(quantile, mean, standard_deviation)
print(paste0("Estimated power: ", estimated_power))

get_estimated_power = function (n, mean, standard_deviation, effect_size, alpha) {
  # quantile = qnorm(0.05)-(effect_size/(standard_deviation/sqrt(n)))
  # estimated_power = 1 - pnorm(quantile)
  power_test = power.t.test(n = n, delta = effect_size, sd = standard_deviation, sig.level = alpha,
             power = NULL,
             type = "two.sample",
             alternative = "one.sided")
  
  return(power_test$power)
}

n = seq(2,100,5) 
x = y = z = NULL
d = seq(0.01,0.35,0.02)
for (delta in d) {
  x = c(x,n)
  y = c(y, rep(delta, NROW(n)))
  z = c(z, sapply(n, get_estimated_power, mean, standard_deviation, delta, alpha))
} 

# sapply(n, get_estimated_power, mean, standard_deviation, effect_size, alpha)
library(scatterplot3d)
library(ggplot2)
library(data.table)

results = data.table(N=x,Power=z, EffectSize=y)
results[Power>0.9, list(Needs_Minimum_N_Subjects=min(N)), by="EffectSize"]

ggplot(data.frame(N=x,Power=z, EffectSize=y), aes(N,Power, color=factor(EffectSize), group=EffectSize)) + geom_line()  + geom_hline(yintercept = 0.9) + ggtitle("Effect size of 0.1 in p_agree - Power analysis for 2 samples one sided t-test") + ylim(c(0,1)) + xlab("N in each group (N/2)")


```

# Try in 3D
Is 3D a better choice to plot these curve ?
```{r}
N = x
effect_size = y
Power = z
threshold = rep(0.9, NROW(Power)) 

s3d = scatterplot3d(N, effect_size, Power, highlight.3d = TRUE, angle = 120,
col.axis = "blue", col.grid = "lightblue", cex.axis = 1.3,
cex.lab = 1.1, main = "Power analysis", pch = 20, mar = c(0.5, 0.3, 0.4, 0.3)) 

# s3d$points3d(effect_size, N, threshold, color = "blue", type = "l", angle = 70)

```
